---
title: "Error_HT-SELEX"
author: "Florian Curvaia"
date: "5/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("reshape2")
library("viridis")
path <- paste(getwd(), "Selex_analysis", sep="/")
```

Warning: This file cannot be run in one single time, because it first create files which then need to be used by the python script Comparison_SELEX.py to generate the files needed for the rest of the code to run.

Chunk which generates the files needed for the python script Comparison_SELEX.py:

```{r, comment="", eval=FALSE, echo=FALSE}
PROX1_nomet <- read.table("PROX1_eDBD_all_nomet_genonets.csv", header=TRUE, sep="\t", nrow=1500)
summary(PROX1_nomet)
write.table(PROX1_nomet, "PROX1_eDBD_best_nomet_genonets.csv", sep="\t", quote = FALSE, row.names = FALSE)
PROX1_met <- read.table("PROX1_eDBD_all_met_genonets.csv", header=TRUE, sep="\t", nrow=1500)
summary(PROX1_met)
write.table(PROX1_met, "PROX1_eDBD_best_met_genonets.csv", sep="\t", quote = FALSE, row.names = FALSE)

PROP1_nomet <- read.table("PROP1_FL_all_nomet_genonets.csv", header=TRUE, sep="\t", nrow=1500)
summary(PROP1_nomet)
write.table(PROP1_nomet, "PROP1_FL_best_nomet_genonets.csv", sep="\t", quote = FALSE, row.names = FALSE)
PROP1_met <- read.table("PROP1_FL_all_met_genonets.csv", header=TRUE, sep="\t", nrow=1500)
summary(PROP1_met)
write.table(PROP1_met, "PROP1_FL_best_met_genonets.csv", sep="\t", quote = FALSE, row.names = FALSE)
```
# PROX1

Generation of the comparison file with python
```{r, comment="", echo=FALSE}
PROX1 <- read.table(paste(path, "PROX1_eDBD_comparison.csv", sep="/"), header=T, sep="\t")
PROX1$Contains_CpG <- as.factor(PROX1$Contains_CpG)
summary(PROX1)
```

Plot with both sequences having CpG and sequences not having CpG
```{r, comment=""}
ggplot(PROX1, aes(x=Score_nomet, y=Score_met, color=Contains_CpG))+geom_point()+geom_abline(slope=1, intercept = 0) + scale_color_viridis(discrete=TRUE, name="Contains CpG", labels=c("No","Yes"))+xlab("Score without methylation")+ylab("Score with methylation")
ggsave(paste(path, "PROX1_scatter.png", sep="/"))
```

Plot with sequences having CpG
```{r, comment=""}
PROX1_CG <- PROX1[PROX1$Contains_CpG == "True",]
PROX1_no_CG <- PROX1[PROX1$Contains_CpG == "False",]
ggplot(PROX1_CG, aes(x=Score_nomet, y=Score_met))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25) + geom_abline(slope=1, intercept = -0.25)
```

Plot with sequences not having CpG
```{r, comment=""}
ggplot(PROX1_no_CG, aes(x=Score_nomet, y=Score_met))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25) + geom_abline(slope=1, intercept = -0.25)
```

Violin plot of distribution of the scores between methylated and non methylated sequences.
```{r, comment=""}
PROX1_mod <- melt(PROX1, measure.vars = c('Score_nomet', 'Score_met'), variable.name="Methylation")
levels(PROX1_mod$Methylation) <- c("No", "Yes")
ggplot(data=PROX1_mod)+geom_violin(aes(Methylation, value, color=Methylation, fill=Methylation))+scale_color_viridis(discrete=TRUE)+scale_fill_viridis(discrete = TRUE, alpha=0.3)+ylab("Score")
ggsave(paste(path, "PROX1_violin.png", sep="/"))
```

Are the non-methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(PROX1$Score_nomet)
```

P-value<<0.05, so we reject the null hypothesis of normal distribution


Are the methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(PROX1$Score_met)
```
P-value<<0.05, so we reject the null hyptohesis of normal distribution


Is there a change between non-methylated and methylated sequences
```{r, comment=""}
wilcox.test(PROX1$Score_nomet, PROX1$Score_met, paired=TRUE)
```
Rejection of null hypothesis of no difference between not methylated and methylated sequences.


Are the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(PROX1$Score_nomet, PROX1$Score_met, paired=TRUE, alternative="less")
```


Are the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(PROX1$Score_nomet, PROX1$Score_met, paired=TRUE, alternative="greater")
```

The results are not really consistent with landscapes' results. Let's try to filter the data by only looking at the sequences which contain at least on CpG dinucleotide.

First, is the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment="", echo=FALSE}
wilcox.test(PROX1_CG$Score_nomet, PROX1_CG$Score_met, paired=TRUE, alternative="less")
```


Then, is the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment="", echo=FALSE}
wilcox.test(PROX1_CG$Score_nomet, PROX1_CG$Score_met, paired=TRUE, alternative="greater")
```

Analyse spread of the scores of sequences which doesn't contain any CpG:
```{r, comment=""}
quantile(PROX1_no_CG$Difference, probs = seq(.1, .9, by = .1))
quantile(PROX1_no_CG$Difference, probs = seq(.05, .95, by = .05))
```

Let's try with absolute difference:
```{r, comment=""}
quantile(abs(PROX1_no_CG$Difference), probs = seq(.1, .9, by = .1))
quantile(abs(PROX1_no_CG$Difference), probs = seq(.05, .95, by = .05))
```

Let's compute the proportion of sequences containing no CpG with a difference inside the 0.25 deviation bound used for now:
```{r, comment=""}
dev_PROX1 <- PROX1_no_CG[abs(PROX1_no_CG$Difference)<=0.25,]
nrow(dev_PROX1)/nrow(PROX1_no_CG)
```

Then let's see if there is a difference in the effect of methylation between sequences with at least one CpG and sequences without any:
```{r, comment=""}
shapiro.test(PROX1_no_CG$Difference)
wilcox.test(PROX1_no_CG$Difference, PROX1_CG$Difference)
```


# PROP1

Generation of the comparison file with python
```{r, comment="", echo=FALSE}
PROP1 <- read.table(paste(path, "PROP1_FL_comparison.csv", sep="/"), header=T, sep="\t")
PROP1$Contains_CpG <- as.factor(PROP1$Contains_CpG)
#PROP1 <- PROP1[PROP1$Score_met>-322,]
summary(PROP1)
```

Plot with both sequences having CpG and sequences not having CpG
```{r, comment=""}
ggplot(PROP1, aes(x=Score_nomet, y=Score_met, color=Contains_CpG))+geom_point()+geom_abline(slope=1, intercept = 0) + scale_color_viridis(discrete=TRUE, name="Contains CpG", labels=c("No","Yes"))+xlab("Score without methylation")+ylab("Score with methylation")
ggsave(paste(path, "PROP1_scatter.png", sep="/"))
```

Plot with sequences having CpG
```{r, comment=""}
PROP1_CG <- PROP1[PROP1$Contains_CpG == "True",]
PROP1_no_CG <- PROP1[PROP1$Contains_CpG == "False",]
ggplot(PROP1_CG, aes(x=Score_nomet, y=Score_met, color=Contains_CpG))+geom_point()+geom_abline(slope=1, intercept = 0) + scale_color_viridis(discrete=TRUE, name="Contains CpG", labels=c("Yes","No"), direction=-1)+xlab("Score without methylation")+ylab("Score with methylation")
ggsave(paste(path, "PROP1_scatter_CG.png", sep="/"))
summary(PROP1_CG)
```

Plot with sequences not having CpG
```{r, comment=""}
ggplot(PROP1_no_CG, aes(x=Score_nomet, y=Score_met))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25) + geom_abline(slope=1, intercept = -0.25)
```

Violin plot of distribution of the scores between methylated and non methylated sequences.
```{r, comment=""}
PROP1_mod <- melt(PROP1, measure.vars = c('Score_nomet', 'Score_met'), variable.name="Methylation")
levels(PROP1_mod$Methylation) <- c("No", "Yes")
ggplot(data=PROP1_mod)+geom_violin(aes(Methylation, value, color=Methylation, fill=Methylation))+scale_color_viridis(discrete=TRUE)+scale_fill_viridis(discrete = TRUE, alpha=0.3)+ylab("Score")
ggsave(paste(path, "PROP1_violin.png", sep="/"))
```


Violin plot of distribution of the scores between methylated and non methylated sequences that contains CG.
```{r, comment=""}
PROP1_CG_mod <- melt(PROP1_CG, measure.vars = c('Score_nomet', 'Score_met'), variable.name="Methylation")
levels(PROP1_CG_mod$Methylation) <- c("No", "Yes")
ggplot(data=PROP1_CG_mod)+geom_violin(aes(Methylation, value, color=Methylation, fill=Methylation))+scale_color_viridis(discrete=TRUE)+scale_fill_viridis(discrete = TRUE, alpha=0.3)+ylab("Score")
ggsave(paste(path, "PROP1_CG_violin.png", sep="/"))
```



Are the non-methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(PROP1$Score_nomet)
```

P-value<<0.05, so we reject the null hypothesis of normal distribution


Are the methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(PROP1$Score_met)
```
P-value<<0.05, so we reject the null hyptohesis of normal distribution


Is there a change between non-methylated and methylated sequences
```{r, comment=""}
wilcox.test(PROP1$Score_nomet, PROP1$Score_met, paired=TRUE)
```
Rejection of null hypothesis of no difference between not methylated and methylated sequences.


Are the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(PROP1$Score_nomet, PROP1$Score_met, paired=TRUE, alternative="less")
```


Are the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(PROP1$Score_nomet, PROP1$Score_met, paired=TRUE, alternative="greater")
```

The results are not really consistent with landscapes' results. Let's try to filter the data by only looking at the sequences which contain at least on CpG dinucleotide.

First, is the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment="", echo=FALSE}
wilcox.test(PROP1_CG$Score_nomet, PROP1_CG$Score_met, paired=TRUE, alternative="less")
```


Then, is the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment="", echo=FALSE}
wilcox.test(PROP1_CG$Score_nomet, PROP1_CG$Score_met, paired=TRUE, alternative="greater")
```

Analyse spread of the scores of sequences which doesn't contain any CpG:
```{r, comment=""}
quantile(PROP1_no_CG$Difference, probs = seq(.1, .9, by = .1))
quantile(PROP1_no_CG$Difference, probs = seq(.05, .95, by = .05))
```

Let's try with absolute difference:
```{r, comment=""}
quantile(abs(PROP1_no_CG$Difference), probs = seq(.1, .9, by = .1))
quantile(abs(PROP1_no_CG$Difference), probs = seq(.05, .95, by = .05))
```

Let's compute the proportion of sequences containing no CpG with a difference inside the 0.25 deviation bound used for now:
```{r, comment=""}
dev_PROP1 <- PROP1_no_CG[abs(PROP1_no_CG$Difference)<=0.25,]
nrow(dev_PROP1)/nrow(PROP1_no_CG)
```

Then let's see if there is a difference in the effect of methylation between sequences with at least one CpG and sequences without any:
```{r, comment=""}
shapiro.test(PROP1_no_CG$Difference)
wilcox.test(PROP1_no_CG$Difference, PROP1_CG$Difference)
```