---
title: "Errors spec-seq"
author: "Florian Curvaia"
date: "27/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("ggplot2")
library("reshape2")
library("viridis")
path <- paste(getwd(), "Sensitivity_analysis_results", sep="/")

```

# AP1

## AP1 sample 1
```{r, comment="", echo=FALSE}
AP1_1 <- read.table(paste(path, "AP1_1/AP1_1_comparison.csv", sep="/"), header=T, sep="\t")
AP1_1$Contains_CpG <- as.factor(AP1_1$Contains_CpG)
summary(AP1_1)
```

Plot of all sequences
```{r, comment=""}
ggplot(AP1_1, aes(x=Score_nomet, y=Score_met, color=Contains_CpG))+geom_point()+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25, linetype="dashed") + geom_abline(slope=1, intercept = -0.25, linetype="dashed")+scale_color_viridis(discrete=TRUE, direction=-1, name="Contains CpG", labels=("Yes"))+xlab("Score without methylation")+ylab("Score with methylation")
ggsave(paste(path, "AP1_1/Plots/AP1_1_scatter.png", sep="/"))
```

Violin plot of distribution of the scores between methylated and non methylated sequences.
```{r, comment=""}
AP1_1_mod <- melt(AP1_1, measure.vars = c('Score_nomet', 'Score_met'), variable.name="Methylation")
levels(AP1_1_mod$Methylation) <- c("No", "Yes")
ggplot(data=AP1_1_mod)+geom_violin(aes(Methylation, value, color=Methylation, fill=Methylation))+scale_color_viridis(discrete=TRUE)+scale_fill_viridis(discrete = TRUE, alpha=0.3)+ylab("Score")
ggsave(paste(path, "AP1_1/Plots/AP1_1_violin.png", sep="/"))
```

Are the non-methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(AP1_1$Score_nomet)

```
P-value<<0.05, so we reject the null hyptohesis of normal distribution

Are the methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(AP1_1$Score_met)
```
P-value<<0.05, so we reject the null hyptohesis of normal distribution

Is there a change between non-methylated and methylated sequences
```{r, comment=""}
wilcox.test(AP1_1$Score_nomet, AP1_1$Score_met, paired=TRUE)
```
Rejection of null hypothesis of no difference between not methylated and methylated sequences

Are the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(AP1_1$Score_nomet, AP1_1$Score_met, paired=TRUE, alternative="less")
```

Are the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(AP1_1$Score_nomet, AP1_1$Score_met, paired=TRUE, alternative="greater")
```

## AP1 sample 2
```{r, comment="", echo=FALSE}
AP1_2 <- read.table(paste(path, "AP1_2/AP1_2_comparison.csv", sep="/"), header=T, sep="\t")
AP1_2$Contains_CpG <- as.factor(AP1_2$Contains_CpG)
#levels(AP1_2$Contains_CpG) <- c(1)
summary(AP1_2)
```

Plot of all sequences
```{r, comment=""}
ggplot(AP1_2, aes(x=Score_nomet, y=Score_met, color=Contains_CpG))+geom_point()+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25, linetype="dashed") + geom_abline(slope=1, intercept = -0.25, linetype="dashed")+scale_color_viridis(discrete=TRUE, direction=-1, name="Contains CpG", labels=("Yes"))+xlab("Score without methylation")+ylab("Score with methylation")
ggsave(paste(path, "AP1_2/Plots/AP1_2_scatter.png", sep="/"))
```

Violin plot of distribution of the scores between methylated and non methylated sequences.
```{r, comment=""}
AP1_2_mod <- melt(AP1_2, measure.vars = c('Score_nomet', 'Score_met'), variable.name="Methylation")
levels(AP1_2_mod$Methylation) <- c("No", "Yes")
ggplot(data=AP1_2_mod)+geom_violin(aes(Methylation, value, color=Methylation, fill=Methylation))+scale_color_viridis(discrete=TRUE)+scale_fill_viridis(discrete = TRUE, alpha=0.3)+ylab("Score")
ggsave(paste(path, "AP1_2/Plots/AP1_2_violin.png", sep="/"))
```

Are the non-methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(AP1_2$Score_nomet)

```
P-value<<0.05, so we reject the null hyptohesis of normal distribution

Are the methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(AP1_2$Score_met)
```
P-value<<0.05, so we reject the null hyptohesis of normal distribution

Is there a change between non-methylated and methylated sequences
```{r, comment=""}
wilcox.test(AP1_2$Score_nomet, AP1_2$Score_met, paired=TRUE)
```
Rejection of null hypothesis of no difference between not methylated and methylated sequences

Are the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(AP1_2$Score_nomet, AP1_2$Score_met, paired=TRUE, alternative="less")
```

Are the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(AP1_2$Score_nomet, AP1_2$Score_met, paired=TRUE, alternative="greater")
```

# CTCF

## CTCF sample 1
```{r, comment=""}
CTCF_1 <- read.table(paste(path, "CTCF_1/CTCF_1_comparison.csv", sep="/"), header=T, sep="\t")
CTCF_1$Contains_CpG <- as.factor(CTCF_1$Contains_CpG)
summary(CTCF_1)
```

Plot with both sequences having CpG and sequences not having CpG
```{r, comment=""}
ggplot(CTCF_1, aes(x=Score_nomet, y=Score_met, color=Contains_CpG))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25, linetype="dashed") + geom_abline(slope=1, intercept = -0.25, linetype="dashed")+scale_color_viridis(discrete=TRUE, name="Contains CpG", labels=c("No","Yes"))+xlab("Score without methylation")+ylab("Score with methylation")
ggsave(paste(path, "CTCF_1/Plots/CTCF_1_scatter.png", sep="/"))
```

Plot with sequences having CpG
```{r, comment=""}
CTCF_1_CG <- CTCF_1[CTCF_1$Contains_CpG == "True",]
CTCF_1_no_CG <- CTCF_1[CTCF_1$Contains_CpG == "False",]
ggplot(CTCF_1_CG, aes(x=Score_nomet, y=Score_met))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25) + geom_abline(slope=1, intercept = -0.25)
```

Plot with sequences not having CpG
```{r, comment=""}
ggplot(CTCF_1_no_CG, aes(x=Score_nomet, y=Score_met))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25) + geom_abline(slope=1, intercept = -0.25)
```

Violin plot of distribution of the scores between methylated and non methylated sequences.
```{r, comment=""}
CTCF_1_mod <- melt(CTCF_1, measure.vars = c('Score_nomet', 'Score_met'), variable.name="Methylation")
levels(CTCF_1_mod$Methylation) <- c("No", "Yes")
ggplot(data=CTCF_1_mod)+geom_violin(aes(Methylation, value, color=Methylation, fill=Methylation))+scale_color_viridis(discrete=TRUE)+scale_fill_viridis(discrete = TRUE, alpha=0.3)+ylab("Score")
ggsave(paste(path, "CTCF_1/Plots/CTCF_1_violin.png", sep="/"))
```

Are the non-methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(CTCF_1$Score_nomet)
```

P-value<<0.05, so we reject the null hypothesis of normal distribution


Are the methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(CTCF_1$Score_met)
```
P-value<<0.05, so we reject the null hyptohesis of normal distribution


Is there a change between non-methylated and methylated sequences
```{r, comment=""}
wilcox.test(CTCF_1$Score_nomet, CTCF_1$Score_met, paired=TRUE)
```
Rejection of null hypothesis of no difference between not methylated and methylated sequences.


Are the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(CTCF_1$Score_nomet, CTCF_1$Score_met, paired=TRUE, alternative="less")
```


Are the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(CTCF_1$Score_nomet, CTCF_1$Score_met, paired=TRUE, alternative="greater")
```

The results are not really consistent with landscapes' results. Let's try to filter the data by only looking at the sequences which contain at least on CpG dinucleotide.

First, is the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment="", echo=FALSE}
wilcox.test(CTCF_1_CG$Score_nomet, CTCF_1_CG$Score_met, paired=TRUE, alternative="less")
```


Then, is the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment="", echo=FALSE}
wilcox.test(CTCF_1_CG$Score_nomet, CTCF_1_CG$Score_met, paired=TRUE, alternative="greater")
```

Analyse spread of the scores of sequences which doesn't contain any CpG:
```{r, comment=""}
quantile(CTCF_1_no_CG$Difference, probs = seq(.1, .9, by = .1))
quantile(CTCF_1_no_CG$Difference, probs = seq(.05, .95, by = .05))
```

Let's try with absolute difference:
```{r, comment=""}
quantile(abs(CTCF_1_no_CG$Difference), probs = seq(.1, .9, by = .1))
quantile(abs(CTCF_1_no_CG$Difference), probs = seq(.05, .95, by = .05))
```

Let's compute the proportion of sequences containing no CpG with a difference inside the 0.25 deviation bound used for now:
```{r, comment=""}
dev_CTCF_1 <- CTCF_1_no_CG[abs(CTCF_1_no_CG$Difference)<=0.25,]
nrow(dev_CTCF_1)/nrow(CTCF_1_no_CG)
```

Then let's see if there is a difference in the effect of methylation between sequences with at least one CpG and sequences without any:
```{r, comment=""}
shapiro.test(CTCF_1_no_CG$Difference)
wilcox.test(CTCF_1_no_CG$Difference, CTCF_1_CG$Difference)
```

## CTCF sample 2
```{r, comment=""}
CTCF_2 <- read.table(paste(path, "CTCF_2/CTCF_2_comparison.csv", sep="/"), header=T, sep="\t")
CTCF_2$Contains_CpG <- as.factor(CTCF_2$Contains_CpG)
summary(CTCF_2)
```

Plot with both sequences having CpG and sequences not having CpG
```{r, comment=""}
ggplot(CTCF_2, aes(x=Score_nomet, y=Score_met, color=Contains_CpG))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25, linetype="dashed") + geom_abline(slope=1, intercept = -0.25, linetype="dashed")+scale_color_viridis(discrete=TRUE, name="Contains CpG", labels=c("No","Yes"))+xlab("Score without methylation")+ylab("Score with methylation")
ggsave(paste(path, "CTCF_2/Plots/CTCF_2_scatter.png", sep="/"))
```

Plot with sequences having CpG
```{r, comment=""}
CTCF_2_CG <- CTCF_2[CTCF_2$Contains_CpG == "True",]
CTCF_2_no_CG <- CTCF_2[CTCF_2$Contains_CpG == "False",]
ggplot(CTCF_2_CG, aes(x=Score_nomet, y=Score_met))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25) + geom_abline(slope=1, intercept = -0.25)
```

Plot with sequences not having CpG
```{r, comment=""}
ggplot(CTCF_2_no_CG, aes(x=Score_nomet, y=Score_met))+geom_point(size=0.25)+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25) + geom_abline(slope=1, intercept = -0.25)
```

Violin plot of distribution of the scores between methylated and non methylated sequences.
```{r, comment=""}
CTCF_2_mod <- melt(CTCF_2, measure.vars = c('Score_nomet', 'Score_met'), variable.name="Methylation")
levels(CTCF_2_mod$Methylation) <- c("No", "Yes")
ggplot(data=CTCF_2_mod)+geom_violin(aes(Methylation, value, color=Methylation, fill=Methylation))+scale_color_viridis(discrete=TRUE)+scale_fill_viridis(discrete = TRUE, alpha=0.3)+ylab("Score")
ggsave(paste(path, "CTCF_2/Plots/CTCF_2_violin.png", sep="/"))
```

Are the non-methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(CTCF_2$Score_nomet)
```

P-value<<0.05, so we reject the null hyptohesis of normal distribution


Are the methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(CTCF_2$Score_met)
```
P-value<<0.05, so we reject the null hyptohesis of normal distribution


Is there a change between non-methylated and methylated sequences
```{r, comment=""}
wilcox.test(CTCF_2$Score_nomet, CTCF_2$Score_met, paired=TRUE)
```
Rejection of null hypothesis of no difference between not methylated and methylated sequences.


Are the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(CTCF_2$Score_nomet, CTCF_2$Score_met, paired=TRUE, alternative="less")
```


Are the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(CTCF_2$Score_nomet, CTCF_2$Score_met, paired=TRUE, alternative="greater")
```

The results are not really consistent with landscapes' results. Let's try to filter the data by only looking at the sequences which contain at least on CpG dinucleotide.

First, is the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment="", echo=FALSE}
wilcox.test(CTCF_2_CG$Score_nomet, CTCF_2_CG$Score_met, paired=TRUE, alternative="less")
```


Then, is the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment="", echo=FALSE}
wilcox.test(CTCF_2_CG$Score_nomet, CTCF_2_CG$Score_met, paired=TRUE, alternative="greater")
```

Analyse spread of the scores of sequences which doesn't contain any CpG:
```{r, comment=""}
quantile(CTCF_2_no_CG$Difference, probs = seq(.1, .9, by = .1))
quantile(CTCF_2_no_CG$Difference, probs = seq(.05, .95, by = .05))
```

Let's try with absolute difference:
```{r, comment=""}
quantile(abs(CTCF_2_no_CG$Difference), probs = seq(.1, .9, by = .1))
quantile(abs(CTCF_2_no_CG$Difference), probs = seq(.05, .95, by = .05))
```

Let's compute the proportion of sequences containing no CpG with a difference inside the 0.25 deviation bound used for now:
```{r, comment=""}
dev_CTCF_2 <- CTCF_2_no_CG[abs(CTCF_2_no_CG$Difference)<=0.25,]
nrow(dev_CTCF_2)/nrow(CTCF_2_no_CG)
```

Then let's see if there is a difference in the effect of methylation between sequences with at least one CpG and sequences without any:
```{r, comment=""}
shapiro.test(CTCF_2_no_CG$Difference)
wilcox.test(CTCF_2_no_CG$Difference, CTCF_2_CG$Difference)
```

Then let's see if there is a difference in the effect of methylation between sequences with at least one CpG and sequences without any:
```{r, comment=""}
shapiro.test(CTCF_2_no_CG$Difference)
wilcox.test(CTCF_2_no_CG$Difference, CTCF_2_CG$Difference)
```

## Comparison of results between sample 1 and sample 2: 

For both samples of CTCF, given the results of the landscape, we would have expected that methylated sequences have a greater score than non-methylated ones. However it doesn't seems to be the case. 
Maybe, methylation only acts in amplifying the scores. For sequences that binds well already when non-methylated, it makes them bind even better, but for sequences that doesn't bind well when non-methylated, maybe then they bind even less when methylated.

# ZFP57
```{r, comment="", echo=FALSE}
ZFP57 <- read.table(paste(path, "ZFP57/ZFP57_comparison.csv", sep="/"), header=T, sep="\t")
ZFP57$Contains_CpG <- as.factor(ZFP57$Contains_CpG)
summary(ZFP57)
```

Plot of all sequences
```{r, comment=""}
ggplot(ZFP57, aes(x=Score_nomet, y=Score_met, color=Contains_CpG))+geom_point()+geom_abline(slope=1, intercept = 0) + geom_abline(slope=1, intercept = 0.25, linetype="dashed") + geom_abline(slope=1, intercept = -0.25, linetype="dashed")+scale_color_viridis(discrete=TRUE, direction=-1, name="Contains CpG", labels=("Yes"))+xlab("Score without methylation")+ylab("Score with methylation")
ggsave(paste(path, "ZFP57/Plots/ZFP57_scatter.png", sep="/"))
```

Violin plot of distribution of the scores between methylated and non methylated sequences.
```{r, comment=""}
ZFP57_mod <- melt(ZFP57, measure.vars = c('Score_nomet', 'Score_met'), variable.name="Methylation")
levels(ZFP57_mod$Methylation) <- c("No", "Yes")
ggplot(data=ZFP57_mod)+geom_violin(aes(Methylation, value, color=Methylation, fill=Methylation))+scale_color_viridis(discrete=TRUE)+scale_fill_viridis(discrete = TRUE, alpha=0.3)+ylab("Score")
ggsave(paste(path, "ZFP57/Plots/ZFP57_violin.png", sep="/"))
```

Are the non-methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(ZFP57$Score_nomet)

```
P-value<<0.05, so we reject the null hyptohesis of normal distribution

Are the methylated scores normally distributed ?
```{r, comment=""}
shapiro.test(ZFP57$Score_met)
```
P-value<<0.05, so we reject the null hyptohesis of normal distribution

Is there a change between non-methylated and methylated sequences
```{r, comment=""}
wilcox.test(ZFP57$Score_nomet, ZFP57$Score_met, paired=TRUE)
```
Rejection of null hypothesis of no difference between not methylated and methylated sequences

Are the scores of methylated sequences greater than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(ZFP57$Score_nomet, ZFP57$Score_met, paired=TRUE, alternative="less")
```

Are the scores of methylated sequences smaller than the one of non-methylated sequences ?
```{r, comment=""}
wilcox.test(ZFP57$Score_nomet, ZFP57$Score_met, paired=TRUE, alternative="greater")
```